---
title: "Lab04PM566"
format: html
editor: visual
---

## Lab 04: Plotting with ggplot()

## Running Code

```{r}
#install.packages("datatable")
library(datasets)
library(leaflet)
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")
unique(met$day)
met

```

### **2. Prepare the data**

```{r}


met_1 <- met[met$temp >-17][elev == 9999.0, elev := NA]
met_1
```

```{r}


met_1$date_variable<-as.Date(paste(met_1$year, met_1$month, met_1$day, sep = "-"))  

met_1$week_number<-data.table::week(as.Date(paste(met_1$year, met_1$month, met_1$day, sep = "-")))
met_1
met_1 <- met_1[met_1$week_number ==31]
unique(met_1$week_number)
```

```{r}

met_avg <- met_1[,.(
  temp     = mean(temp,na.rm=TRUE),
  rh       = mean(rh,na.rm=TRUE),
  wind.sp  = mean(wind.sp,na.rm=TRUE),
  vis.dist = mean(vis.dist,na.rm=TRUE),
  lat      = mean(lat),
  lon      = mean(lon), 
  elev     = mean(elev,na.rm=TRUE),
  dew.point =mean(dew.point, na.rm=TRUE)
), by=c("USAFID", "day")]

met_avg
```

```{r}

met_avg$regionlat <- cut(met_avg$lat,
                   breaks=c(24.55525, 39.71 ,48.941),
                   labels=c('N','S'))
met_avg$regionlon <-cut(met_avg$lon,
                   breaks=c(-124.29, -98.00 ,-68.313),
                   labels=c('E','W'))

met_avg$region<- paste(met_avg$regionlat, met_avg$regionlon, sep = "")
met_avg


                                                                                               
```

```{r}
library(ggplot2)
library(magrittr)
met_avg

met_avg[!is.na(wind.sp)] %>% 
  ggplot()+
  geom_boxplot(mapping=aes(x=1, y=dew.point, fill=region))

```

DESCRIBE THE PLOT:

### **3. Use `geom_violin` to examine the wind speed and dew point by region**

```{r}

met_avg[!is.na(wind.sp)] %>% 
  ggplot()+
  geom_violin(mapping=aes(x=1, y=dew.point,fill=region))+
  labs(title= "Wind Speed and Dewpoint by Region", x="Wind Speed",y="Dewpoint")

```

Describe the plot: I see that there is more of a spread in dew point for SE and NE regions. In terms of wind speed, the data is more spread out for NW and SW.

### **4. Use `geom_jitter` with `stat_smooth` to examine the association between dew point and wind speed by region**

```{r}

met_avg[!is.na(dew.point) & !is.na(wind.sp) ] %>%
  ggplot() + 
  geom_point(mapping = aes(x = dew.point, y = wind.sp, color = region)) +geom_smooth(mapping = aes(x = dew.point, y = wind.sp, linetype = region))+
  labs(title= "Dew Point and Wind Speed by Region", x="Dew Point",y="Windspeed")

```

I think there is probably a better way to visualize this since the legend doesn't seem to give lines that are different than each other. I noticed that the association between the two lines in the middle are closer than the other two lines.

### **5. Use `geom_bar` to create barplots of the weather stations by elevation category colored by region (Working)**

```{r}



met_avg[, elev_cat := ifelse(elev > 252, "high", "low")]
met_avg[, vis_cat  := cut(
  x      = vis.dist,
  breaks = c(0, 1000, 6000, 10000, Inf),
  labels = c("fog", "mist", "haze", "clear"),
  right  = FALSE
)]


met_avg[!is.na(vis_cat)] %>%
  ggplot() + 
  geom_bar(mapping = aes(x =elev_cat, fill=region),position = "dodge") + scale_fill_brewer()+
  labs(title= "Elevation Categories by Region", x="Elevation Categories",y="Count")



```

Describe the plot. I had a hard time figuring out how to include USAFID since I would get an error when I made it my X and Y value. However, this graph shows that elevation categories by region. We see that NW is mosly categorized by low elevation. SE has less elevation categories that are ocnsidered low.

### **6. Use `stat_summary` to examine mean dew point and wind speed by region with standard deviation error bars (Working)**

```{r}




 met_avg[!is.na(dew.point)]%>%
  ggplot() + 
    stat_summary(mapping = aes(x = wind.sp, y = region),
    fun.min = min,
    fun.max = max,
    fun = median,
    fun.data =mean_sdl) + stat_summary(mapping = aes(x = wind.sp, y = region),geom="errorbar")+
  labs(title= "Mean Dew Point by Region", x="Mean Dew Point ",y="Region")



 met_avg[!is.na(wind.sp)]%>%
  ggplot() + 
    stat_summary(mapping = aes(x = wind.sp, y = region),
    fun.min = min,
    fun.max = max,
    fun = median,
    fun.data =mean_sdl) + stat_summary(mapping = aes(x = wind.sp, y = region),geom="errorbar")+
  labs(title= "Wind Speed by Region", x="Wind Speed",y="Region")


```

The mean between NE and SW seem to be closer to each other then other means.

The highest mean value in dew point is NE.

The highest mean value in wind speed is NE.

### **7. Make a map showing the spatial trend in relative humidity in the US (Working)**

```{r}
# Make sure to remove NA
met_avg <- met_avg[!is.na(rh)]
met_avg

#Make a color palette with custom colors
rh.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)
rh.pal


#Use Leaflet  a
RH_map <- leaflet(met_avg) %>%
  
  addProviderTiles('CartoDB.Positron') %>% 
  
  addCircles(lat = ~lat, lng=~lon,label = ~paste0(round(rh,2), 'RH'),color = ~ rh.pal(rh),opacity = 1, fillOpacity = 1, radius = 500)  %>% addLegend('bottomleft', pal=rh.pal, values=met_avg$rh,
          title='Relative Humidity', opacity=1)# %>% addMarkers(~lon, ~lat, popup = ~rank(-rh) <= 10, label = ~rank(-rh) <= 10)
RH_map

```

The more east you go, the more humid it gets. There is more data in the eastern part of the United States compared to the west side.

```         
```

### **8. Use a ggplot extension**

```{r}

library(ggplot2)
library(patchwork)

met_avg

p1 <- ggplot(met_avg) + geom_point(aes(dew.point,wind.sp))
p2 <- ggplot(met_avg) + geom_boxplot(aes(region, wind.sp, group = region))

p1 + p2


p3 <- ggplot(met_avg) + geom_smooth(aes(wind.sp,rh))
p4 <- ggplot(met_avg) + geom_bar(aes(region))

(p1 | p2 | p3) /
      p4

```
